Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago

### 2024-07-10 manuscript.Rmd
acknowledgments/keywords and updating renv

---
title: 'Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago'
author:
  - Kaleonani K.C. Hurley:
      institute:
        - himb
      equal_contributor: yes
  - Mykle L. Hoban:
      institute:
        - himb
      equal_contributor: yes
      correspondence: yes
      email: "mhoban@hawaii.edu"
  - Kerry Reardon:
      institute: noaa
  - Derek J. Skillings:
      institute: 
        - uncg
  - Molly A. Timmers:
      institute:
        - himb
        - natgeo
  - Robert J. Toonen:
      institute: himb
institute:
  - himb: Hawai&#x02BB;i Institute of Marine Biology, University of Hawai&#x02BB;i, Mānoa
  - noaa: National Oceanic and Atmospheric Administration
  - uncg: 'University of North Carolina, Greensboro'
  - natgeo: National Geographic Pristine Seas
abstract:
  An inability to consistently resolve the environmental and anthropogenic factors that shape coral reefs impacts our understanding of how those factors interact and how coral reef communities may respond to change. Logistical challenges have additionally limited the taxonomic resolution and depth ranges investigated, with the focus largely on fishes and corals in the shallowest ~30 m of coral reefs. Here, we examine the effects of a suite of environmental, ecological, and anthropogenic factors on the assemblage structure of brachyuran crabs recovered from Autonomous Reef Monitoring Structures on shallow reefs (8--17 m) across the ~2,500 km span of the Hawaiian Archipelago. We compare these assemblages with those sampled along a depth gradient (12--90 m) on O&#x02BB;ahu. We found 2,284 individuals representing 103 (morpho)species from 22 families. Shallow assemblages were significantly different between the Main and Northwest Hawaiian Islands and significantly associated with human impacts, chlorophyll-A, sea-surface temperature, slope, potential larval connectivity, and depth. However, we found more variation among mesophotic samples from O&#x02BB;ahu (30--90 m) than among shallow samples across the entire archipelago, suggesting that 90 m of depth is a stronger determinant of brachyuran assemblage structure than the latitudinal, environmental, or anthropogenic gradients across the Hawaiian Islands.
bibliography: references/citations.json
csl: references/style.csl
output:
  word_document:
    pandoc_args:
      - "--toc-depth=2"
      - "--filter=resources/pandoc-crossref"
      - "--lua-filter=resources/scholarly-metadata.lua"
      - "--lua-filter=resources/author-info-blocks.lua"
      - "--reference-doc=resources/ref.docx"
    keep_md: false
tblPrefix: table
figPrefix:
  - "figure"
  - "figures"
chapters: True
chaptersDepth: 1
chapDelim: ""
header-includes:
  - \usepackage{bm}
  - \usepackage{upgreek}
---
<!--word count: Proceedings B publishes articles up to 12 pages, including figures and tables, with no more than 10 000 words. The word count includes the title page, author names, abstract, figure legends, tables and the reference list. As a rough guide, 10 000 words = 12 pages without additional figures; 850 words = 1 page and 2-4 small figures = 1 page. For example, an acceptable paper could be made up of 9 000 words and 4 small figures or 8500 words, 2 large figures and 4 small figures.-->
### Keywords {-}
<div custom-style="mono">ARMs, mesophotic coral ecosystems, community structure, dispersal, Hawai&#x02BB;i, brachyuran crabs</div>

@@

# Acknowledgements
We thank the Coral Reef Conservation Program for supporting the deployment and recovery of the shallow ARMS. We thank the UH Mānoa Dive Safety Office (D. Pence, K. Stender, J. Jones, and C.J. Bradley) for installation and retrieval of the mesophotic ARMS. We thank Joshua Copus and Kevin Flanagan for additional assistance with retrieval and processing of mesophotic ARMS units. We thank the ToBo laboratory members, particularly M. Iacchei, C. Ka&#x02BB;apu-Lyons, and Z. Hee, for discussion, support, and their efforts in the disassembly and sorting of ARMS samples on retrieval. This is contribution #xxx from the Hawai&#x02BB;i Institute of Marine Biology and #xxx from the School of Earth Science and Technology at the University of Hawai&#x02BB;i.

# Funding
This work was funded through a combination of grants including NSF OCE 12-60169, NSF GRFP DGE-1329626, and NSF-2048457. In addition, the NOAA Coral Reef Conservation Program funded the Pacific Reef Assessment and Monitoring Program, which included the shallow ARMS deployment in the Hawaiian Archipelago.

>>>>

---
title: 'Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago'
author:
  - Kaleonani K.C. Hurley:
      institute:
        - himb
      equal_contributor: yes
  - Mykle L. Hoban:
      institute:
        - himb
      equal_contributor: yes
      correspondence: yes
      email: "mhoban@hawaii.edu"
  - Kerry Reardon:
      institute: noaa
  - Derek J. Skillings:
      institute: 
        - uncg
  - Molly A. Timmers:
      institute:
        - himb
        - natgeo
  - Robert J. Toonen:
      institute: himb
institute:
  - himb: Hawai&#x02BB;i Institute of Marine Biology, University of Hawai&#x02BB;i, Mānoa
  - noaa: National Oceanic and Atmospheric Administration
  - uncg: 'University of North Carolina, Greensboro'
  - natgeo: National Geographic Pristine Seas
abstract:
  An inability to consistently resolve the environmental and anthropogenic factors that shape coral reefs impacts our understanding of how those factors interact and how coral reef communities may respond to change. Logistical challenges have additionally limited the taxonomic resolution and depth ranges investigated, with the focus largely on fishes and corals in the shallowest ~30 m of coral reefs. Here, we examine the effects of a suite of environmental, ecological, and anthropogenic factors on the assemblage structure of brachyuran crabs recovered from Autonomous Reef Monitoring Structures on shallow reefs (8--17 m) across the ~2,500 km span of the Hawaiian Archipelago. We compare these assemblages with those sampled along a depth gradient (12--90 m) on O&#x02BB;ahu. We found 2,284 individuals representing 103 (morpho)species from 22 families. Shallow assemblages were significantly different between the Main and Northwest Hawaiian Islands and significantly associated with human impacts, chlorophyll-A, sea-surface temperature, slope, potential larval connectivity, and depth. However, we found more variation among mesophotic samples from O&#x02BB;ahu (30--90 m) than among shallow samples across the entire archipelago, suggesting that 90 m of depth is a stronger determinant of brachyuran assemblage structure than the latitudinal, environmental, or anthropogenic gradients across the Hawaiian Islands.
bibliography: references/citations.json
csl: references/style.csl
output:
  word_document:
    pandoc_args:
      - "--toc-depth=2"
      - "--filter=resources/pandoc-crossref"
      - "--lua-filter=resources/scholarly-metadata.lua"
      - "--lua-filter=resources/author-info-blocks.lua"
      - "--reference-doc=resources/ref.docx"
    keep_md: false
tblPrefix: table
figPrefix:
  - "figure"
  - "figures"
chapters: True
chaptersDepth: 1
chapDelim: ""
header-includes:
  - \usepackage{bm}
  - \usepackage{upgreek}
---
<!--word count: Proceedings B publishes articles up to 12 pages, including figures and tables, with no more than 10 000 words. The word count includes the title page, author names, abstract, figure legends, tables and the reference list. As a rough guide, 10 000 words = 12 pages without additional figures; 850 words = 1 page and 2-4 small figures = 1 page. For example, an acceptable paper could be made up of 9 000 words and 4 small figures or 8500 words, 2 large figures and 4 small figures.-->
### Keywords {-}
<div custom-style="mono">Mesophotic coral ecosystems, community structure, depth gradient, environmental drivers, Hawaii, Autonomous Reef Monitoring Structures</div>

@@

# Acknowledgements
We thank the NOAA Coral Reef Conservation Program for supporting the deployment and recovery of the shallow ARMS. We thank the UH Mānoa Dive Safety Office (D. Pence, K. Stender, J. Jones, and C.J. Bradley) for installation and retrieval of the mesophotic ARMS. We specially thank K. Flanagan from the UH dive program, whose support was instrumental in the development of the mesophotic ARMS project, and J. Copus, who provided expert mesophotic field support; both were taken too soon. May their memories be a blessing. We thank the members of the ToBo lab, particularly M. Iacchei, C. Ka&#x02BB;apu-Lyons, and Z. Hee, for discussion, support, and their efforts in the disassembly and sorting of ARMS samples on retrieval. This is contribution #xxx from the Hawai&#x02BB;i Institute of Marine Biology and #xxx from the School of Ocean and Earth Science and Technology at the University of Hawai&#x02BB;i.

# Funding
This work was funded through a combination of grants including NSF OCE-1260169, NSF GRFP DGE-1329626, and NSF OCE-2048457. In addition, the NOAA Coral Reef Conservation Program funded the Pacific Reef Assessment and Monitoring Program, which included the shallow ARMS deployment in the Hawaiian Archipelago.
<!--
Specimens used in this study were collected under Hawai&#x02BB;i Division of Aquatic Resources Special Activity Permit numbers SAP 2013-47, SAP 2012-63, SAP 2011-01, and SAP 2008-99.
-->
